## assignment Histogram 

lesson 2 (cont.) professor's solution

2020-02-13

Using the *stream* package in R, keep current sample histogram of a 1-dimensional data stream using different window models.

**keep sample histogram**

defenitions
```{r}
library(stream)

#par(mfrow=c(2,2))

# definitions
itera <- 2000
lbound <- -3
ubound <- 8
nbins <- 10
mu1 <- 0
mu2 <- 5

```

iteration
```{r}
# initialization
set.seed(523)
stream1 <- DSD_Gaussians(k=1, d=1, mu=mu1, sigma=2)
stream2 <- DSD_Gaussians(k=1, d=1, mu=mu2, sigma=2)
splits <- seq(from=lbound, to=ubound, by=(ubound-lbound)/nbins)
splits <- c(splits, +Inf) # extra split for comparison
bins <- rep(0, length(splits))
names(bins) <- paste("<=",splits,sep="")
plot(NULL, xlab="Iteration",ylab="Value", xlim=c(0,itera), 
     ylim=c(lbound,ubound), main="Sample Distribution Heat Map")

cat("\n*************************\n", "Sample Distribution","\n")

# iterative step
for (i in 1:itera){
  xi <- ifelse(i<=1000,
               get_points(stream1)[1,1],
               get_points(stream2)[1,1])
  
  bin <- which(xi <= splits)[1]
  
  bins[bin] <- bins[bin] + 1
  
  sapply(2:(length(splits)-1), function(u){
    upper <- splits[u]
    lower <- splits[u-1]
    lines(x=c(i,i), y=c(lower,upper), lwd=2, col=gray(1 - bins[u] / sum(bins)))
  })
  if (i %% 500 == 0){
    cat("\n ### Iteration",i, "\n")
    print(bins)
  }
}
```

**keep sample histogram over a sliding window**

definitions and iteration
```{r}
# definitions
w <- 500

# initialization
set.seed(523)
stream1 <- DSD_Gaussians(k=1, d=1, mu=mu1, sigma=2)
stream2 <- DSD_Gaussians(k=1, d=1, mu=mu2, sigma=2)
splits <- seq(from=lbound, to=ubound, by=(ubound-lbound)/nbins)
splits <- c(splits, +Inf) # extra split for comparison
bins <- rep(0, length(splits))
names(bins) <- paste("<=",splits,sep="")
window <- NULL

plot(NULL, xlab="Iteration",ylab="Value", xlim=c(0,itera), 
     ylim=c(lbound,ubound), main="Sliding Distribution Heat Map")

cat("\n*************************\n", "Sliding Distribution","\n")

# iterative step
for (i in 1:itera){
  xi <- ifelse(i<=1000,
               get_points(stream1)[1,1],
               get_points(stream2)[1,1])
  
  bin <- which(xi <= splits)[1]
  
  bins[bin] <- bins[bin] + 1
  
  window <- c(window, xi)
  
  if (length(window) >  w) {
    oldbin <- which(window[1] <= splits)[1]
    bins[oldbin] <- bins[oldbin] - 1
    window <- window[-1]
  }
  
  sapply(2:(length(splits)-1), function(u){
    upper <- splits[u]
    lower <- splits[u-1]
    lines(x=c(i,i), y=c(lower,upper), lwd=2, col=gray(1 - bins[u] / sum(bins)))
  })
  if (i %% 500 == 0){
    cat("\n ### Iteration",i, "\n")
    print(bins)
  }
}
```

**keep sample histogram over a alpha weighted sliding window**

definitions
```{r}
w <- 500
eps <- 0.05
alpha <- eps^(1/w)

```

iteration
```{r}
# initialization
set.seed(523)
stream1 <- DSD_Gaussians(k=1, d=1, mu=mu1, sigma=2)
stream2 <- DSD_Gaussians(k=1, d=1, mu=mu2, sigma=2)
splits <- seq(from=lbound, to=ubound, by=(ubound-lbound)/nbins)
splits <- c(splits, +Inf) # extra split for comparison
bins <- rep(0, length(splits))
names(bins) <- paste("<=",splits,sep="")
window <- NULL

plot(NULL, xlab="Iteration",ylab="Value", xlim=c(0,itera), 
     ylim=c(lbound,ubound), main="Weighted Distribution Heat Map")

cat("\n*************************\n", "Weighted Distribution","\n")

# iterative step
for (i in 1:itera){
  xi <- ifelse(i<=1000,
               get_points(stream1)[1,1],
               get_points(stream2)[1,1])
  
  bin <- which(xi <= splits)[1]
  
  bins <- bins * alpha
  bins[bin] <- bins[bin] + 1
  
  window <- c(window, xi)
  
  if (length(window) >  w) {
    oldbin <- which(window[1] <= splits)[1]
    bins[oldbin] <- bins[oldbin] - alpha^w
    window <- window[-1]
  }
  
  sapply(2:(length(splits)-1), function(u){
    upper <- splits[u]
    lower <- splits[u-1]
    lines(x=c(i,i), y=c(lower,upper), lwd=2, col=gray(1 - bins[u] / sum(bins)))
  })
  if (i %% 500 == 0){
    cat("\n ### Iteration",i, "\n")
    print(bins)
  }
}
```

**keep sample histogram over a alpha fading window**

definitions
```{r}

w <- 500
eps <- 0.05
alpha <- eps^(1/w)

```

iteration
```{r}
# initialization
set.seed(523)
stream1 <- DSD_Gaussians(k=1, d=1, mu=mu1, sigma=2)
stream2 <- DSD_Gaussians(k=1, d=1, mu=mu2, sigma=2)
splits <- seq(from=lbound, to=ubound, by=(ubound-lbound)/nbins)
splits <- c(splits, +Inf) # extra split for comparison
bins <- rep(0, length(splits))
names(bins) <- paste("<=",splits,sep="")
plot(NULL, xlab="Iteration",ylab="Value", xlim=c(0,itera), 
     ylim=c(lbound,ubound), main="Fading Distribution Heat Map")

cat("\n*************************\n", "Fading Distribution","\n")
# iterative step
for (i in 1:itera){
  xi <- ifelse(i<=1000,
               get_points(stream1)[1,1],
               get_points(stream2)[1,1])
  
  bin <- which(xi <= splits)[1]
  
  bins <- bins * alpha
  bins[bin] <- bins[bin] + 1
  
  sapply(2:(length(splits)-1), function(u){
    upper <- splits[u]
    lower <- splits[u-1]
    lines(x=c(i,i), y=c(lower,upper), lwd=2, col=gray(1 - bins[u] / sum(bins)))
  })
  if (i %% 500 == 0){
    cat("\n ### Iteration",i, "\n")
    print(bins)
  }
}
```

