## assignment Matrix 

Assignment 1 (deadline Feb 17)

2020-02-15

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
```

Using R, keep counters of equal-width grid-cells (base counters for micro-cluster definitions) of a 2-dimensional continuous data stream using different window models (landmark, sliding, weighted, fading).

The code should work with an evolving stream from a single record of the PhysioNet Challenge https://physionetchallenges.github.io/2020/

This training set consists of 6,877 (male: 3,699; female: 3,178) 12-ECG recordings lasting from 6 seconds to 60 seconds. Each recording was sampled at 500 Hz. All data is provided in WFDB format with a MATLAB v4 file and a header containing patient sex, age, and diagnosis (Dx) information at the end of the header file. The code should be applicable to one 12-dimensional record file.


```{r}
library(stream)
library(R.matlab)

# after unzip the files, I took the first file 
fA1<-readMat("2.UploadedData/A0001.mat")$val
fA1<-t(fA1)
```

I selected two columns (4 and 5) corresponding to the readings and make the streams
```{r}
# select columns
a<-4
b<-5
# data stream
strA1<-DSD_Memory(fA1[,c(a,b)], loop = T)
```

definitions
```{r}
set.seed(1)
reset_stream(strA1)
# number of iterations
itera <- 2000
#defining bounds for matrix
lbound <- min(strA1[[2]])
ubound <- max(strA1[[2]])
nbins <- 10

# first we create two equal size vectors
splits <- seq(from=lbound, to=ubound, by=(ubound-lbound)/nbins)
splits <- c(splits)[-1] # drop first value
binsA <- rep(0, length(splits))
names(binsA) <- paste("<=",splits,sep="")

binsB <- rep(0, length(splits))
names(binsB) <- paste("<=",splits,sep="")

# and then a matrix 10x10
mx<-matrix(0,10,10)
rownames(mx)<-colnames(mx)<-names(binsA)

```

### keep sample matrix

iteration
```{r}
# iterative step
for (i in 1:itera){
  
  ai <- get_points(strA1)[1,1]
  bi <- get_points(strA1)[1,2]
  
  binA <- which(ai <= splits)[1]
  binsA[binA] <- binsA[binA] + 1
  
  binB <- which(bi <= splits)[1]
  binsB[binB] <- binsB[binB] + 1
  
  mx[binA,binB]<-mx[binA,binB] + 1 # with the vectors positions, we update the matrix counts
  
  # printing the two vectors just for control
  if (i %% 500 == 0){
    cat("\n ### Iteration",i, "\n")
    print(binsA)
    print(binsB)
  }
}
```

heat map with the values of the matrix

```{r}
heatmap(mx, Colv = NA, Rowv = NA, scale="none", symm = F, revC = T,
        main = "Sample Distribution Heat Map")
```

### keep sample matrix over a sliding window

definitions and initialization
```{r}
set.seed(1)
reset_stream(strA1)
w <- 500

binsA <- rep(0, length(splits))
names(binsA) <- paste("<=",splits,sep="")

binsB <- rep(0, length(splits))
names(binsB) <- paste("<=",splits,sep="")

mx<-matrix(0,10,10)
rownames(mx)<-colnames(mx)<-names(binsA)

windowA <- windowB <- NULL
```

iteration
```{r}
# iterative step
for (i in 1:itera){
  ai <- get_points(strA1)[1,1]
  bi <- get_points(strA1)[1,2]
  
  binA <- which(ai <= splits)[1]
  binsA[binA] <- binsA[binA] + 1
  
  binB <- which(bi <= splits)[1]
  binsB[binB] <- binsB[binB] + 1
  
  mx[binA,binB]<-mx[binA,binB] + 1
  
  windowA <- c(windowA, ai) # defining two windows
  windowB <- c(windowB, bi) # one for each stream
  
    if (length(windowA) >  w) {
    oldbinA <- which(windowA[1] <= splits)[1]
    binsA[oldbinA] <- binsA[oldbinA] - 1
    windowA <- windowA[-1]
    
    oldbinB <- which(windowB[1] <= splits)[1]
    binsB[oldbinB] <- binsB[oldbinB] - 1
    windowB <- windowB[-1]
    
    mx[oldbinA,oldbinB]<-mx[oldbinA,oldbinB] - 1 # droping the first value from the window
    }
  
  # printing the two vectors just for control
  if (i %% 500 == 0){
    cat("\n ### Iteration",i, "\n")
    print(binsA)
    print(binsB)
  }
}
```

heat map with the values of the matrix
```{r}
heatmap(mx, Colv = NA, Rowv = NA, scale="none", symm = F, revC = T,
        main = "Sliding Distribution Heat Map")
```

### keep sample matrix over a alpha weighted sliding window

definitions and initialization
```{r}
set.seed(1)
reset_stream(strA1)

w <- 500
eps <- 0.05
alpha <- eps^(1/w) # defining an alpha value

# initialization
set.seed(1)
binsA <- rep(0, length(splits))
names(binsA) <- paste("<=",splits,sep="")

binsB <- rep(0, length(splits))
names(binsB) <- paste("<=",splits,sep="")

mx<-matrix(0,10,10)
rownames(mx)<-colnames(mx)<-names(binsA)

windowA <- windowB <- NULL
```

iteration
```{r}
# iterative step
for (i in 1:itera){
  ai <- get_points(strA1)[1,1]
  bi <- get_points(strA1)[1,2]
  
  binA <- which(ai <= splits)[1]
  binsA <-binsA * alpha
  binsA[binA] <- binsA[binA] + 1
  
  binB <- which(bi <= splits)[1]
  binsB <-binsB * alpha
  binsB[binB] <- binsB[binB] + 1
  
  mx<-mx * alpha #multiplying the matrix by alpha before updating it 
  mx[binA,binB]<-mx[binA,binB] + 1 # updating the matrix with a new count
  
  windowA <- c(windowA, ai)
  windowB <- c(windowB, bi)
  
  if (length(window) >  w) {
    oldbinA <- which(windowA[1] <= splits)[1]
    binsA[oldbinA] <- binsA[oldbinA] - 1
    windowA <- windowA[-1]
    
    oldbinB <- which(windowB[1] <= splits)[1]
    binsB[oldbinB] <- binsB[oldbinB] - 1
    windowB <- windowB[-1]
    
    mx[oldbinA,oldbinB]<-mx[oldbinA,oldbinB] - alpha^w
  }
  
    if (i %% 500 == 0){
    cat("\n ### Iteration",i, "\n")
    print(binsA)
    print(binsB)
  }
}
```

heat map with the values of the matrix
```{r}
heatmap(mx, Colv = NA, Rowv = NA, scale="none", symm = F, revC = T,
        main = "Weighted Distribution")
```

### keep sample matrix over a alpha fading window

definitions and initialization
```{r}
set.seed(1)
reset_stream(strA1)

w <- 500
eps <- 0.05
alpha <- eps^(1/w)

binsA <- rep(0, length(splits))
names(binsA) <- paste("<=",splits,sep="")

binsB <- rep(0, length(splits))
names(binsB) <- paste("<=",splits,sep="")

mx<-matrix(0,10,10)
rownames(mx)<-colnames(mx)<-names(binsA)
```

iteration
```{r}
# iterative step
for (i in 1:itera){
  ai <- get_points(strA1)[1,1]
  bi <- get_points(strA1)[1,2]
  
  binA <- which(ai <= splits)[1]
  binsA[binA] <- binsA[binA] + 1
  
  binB <- which(bi <= splits)[1]
  binsB[binB] <- binsB[binB] + 1
  
  mx<-mx * alpha # multiplying the matrix by alpha before updating it
  mx[binA,binB]<-mx[binA,binB] + 1 # updating the matix
  
    if (i %% 500 == 0){
    cat("\n ### Iteration",i, "\n")
    print(binsA)
    print(binsB)
  }
}
  
```

heat map with the values of the matrix
```{r}

heatmap(mx, Colv = NA, Rowv = NA, scale="none", symm = F, revC = T,
        main = "Fading Distribution")
```


### Bonus
for a diferent solution

https://rpubs.com/franzbischoff/hida_assign1
