## Lesson 3

2020-02-17

**Classe Exercises** (almost all code given by the professor)

```{r }
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

### Kernel density estimation
```{r}
data(faithful)
hist(faithful$waiting, density = 10,
     main="wainting time to next eruption",
     xlab="minutes", freq=F)
lines(density(faithful$waiting),lwd=2,col=4)
lines(density(faithful$waiting, bw = 5),lwd=2,col=5) # defining a banwith = 5 in density function
lines(density(faithful$waiting, bw = 10),lwd=2,col=6) # defining a banwith = 10 in density function

```

### Conditional density
With data from Coimbra Breast Cancer (Glucose | Classification)
```{r}
breast<-read.csv("2.UploadedData/dataR2.csv")

hist(breast$Glucose, density = 10,
     main="Glucose level",
     xlab="", freq=F, ylim = c(0,0.05))
lines(density(breast$Glucose),
      lwd=2,col="black")
lines(density(subset(breast,Classification == 1)$Glucose),
      lwd=2,col="green")
lines(density(subset(breast,Classification == 2)$Glucose),
      lwd=2,col="red")
legend(130,0.03, 
       legend=c("P(Glucose)","P(Glucose|Control)","P(Glucose|Patient)"), 
       col=c("black","green","red"),
       lty = 1,
       cex = 0.8
       )
```

### Expectation Maximization for the two latent variables
Example with Age distribution in the breast cancer data set
```{r}
# read data
breast<-read.csv("2.UploadedData/dataR2.csv")

# Expectation Maximization for the two latent variables (clusters)
m<-2 # the total number of clusters (m)
set.seed(0)
param.mean<-sample(breast$Age,m) # initial parameters
param.sd<-rep(sd(breast$Age)/m,m) # initial parameters

# E-step (Expectation): that estimates the probability of each point belonging to each cluster
xx<-breast$Age

## definition of M-step method
alternative <- 1

## iteraction process
itera <-10 # maximum number of iterations
eps<-0.01 # minimum error to convergence
for (k in 1:itera) {
densities<-sapply(1:m, 
                  function(i){dnorm(xx,
                                    mean=param.mean[i],
                                    sd=param.sd[i])})
relevance<-t(apply(densities,1,function(l){l/sum(l)}))

attrib<-t(apply(relevance,1,
                function(l){replace(rep(0,m),which.max(l)[1],1)}))

# Re-estimates the parameters of the probability distribution of each class
if (alternative == 1) {
## M-step (alternative 1: crisp)
  param.mean.new<-sapply(1:m,
                         function(j){mean(breast$Age[attrib[,j]==1])})
  param.sd.new<-sapply(1:m,
                       function(j){sqrt(sum(relevance[,j]*(breast$Age-param.mean[j])^2)/sum(relevance[,j]))})
} else {
## M-step (alternative 2: soft)
  param.mean.new<-sapply(1:m,
                         function(j){sum(breast$Age*relevance[,j])/sum(relevance[,j])})
  param.sd.new<-sapply(1:m,
                       function(j){sqrt(sum(relevance[,j]*(breast$Age-param.mean[j])^2)/sum(relevance[,j]))})
  
}

# convergence the maximum number of iterations (k).
cat("iteration", itera, "evolution:", 
    dif<-sum(abs(param.mean.new - param.mean)), "\n")
if(dif < eps) break # the accepted error to converge (e)
param.mean<-param.mean.new
param.sd<-param.sd.new
}
```

Print the two new density lines with a large banwith
```{r}
# plot histogram and density function for Age
hist(breast$Age, density = 20,
     main="Glucose level",
     xlab="", freq=F, ylim = c(0,0.03))
lines(density(breast$Age),
      lwd=2,col="black")

set.seed(0)
lines(density(rnorm(100, param.mean[1], param.sd[1]), bw = 12),lwd=2,col=5)
lines(density(rnorm(100, param.mean[2], param.sd[2]), bw = 12),lwd=2,col=4)
```

### Data Generation
Comparing the distribution from data generated from 3 different ways
```{r}
## length of data set generation
n<-100

# original sample data set
sample<-breast$Age

# bootstrapped sample data set
sample.obs<-sample(breast$Age, size=n, replace = TRUE)

# gaussian generated data set
sample.simple<-rnorm(n,mean=mean(breast$Age), sd=sd(breast$Age))

# EM model generated data set
classes<-apply(attrib, 1, which.max) # using attrib define previously, to calculate the number of observations in each class
class.prop<-prop.table(table(classes)) # calculate percentage of each class
set.seed(523)
sample.classes<-sample(1:m, size=n, replace=TRUE, prob=class.prop)
sample.prop<-prop.table(table(sample.classes))
sample.values<-round(unlist(sapply(1:m,
                                   function(j){
                                     rnorm(n*sample.prop[j], # generating number with the proportions of the two classes we identified previously
                                           mean=param.mean[j], # parameters defined previously
                                           sd=param.sd[j]) # parameters defined previously
                                   })),0)
```

Printing the distribution of the 4 data samples
```{r}
par(mfrow=c(2,2))

hist(breast$Age, freq = F, main = "Observed data")
lines(density(breast$Age))

hist(sample.obs, freq = F, main = "data points sampled from observed data")
lines(density(sample.obs))

hist(sample.simple, freq = F, main = "data points generated from sample distribution")
lines(density(sample.simple))

hist(sample.values, freq = F, main = "data points generated from EM (m=2)")
lines(density(sample.values))
```

