## assignment Glucose 

Assignment from lesson 3

2020-02-26

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
```

Using only the base and stats packages in R, define the iterative steps of ExpectationMaximization applied to the Glucose variable of the Breast Cancer Coimbra data set,
searching for 2 or more latent classes.

### Defining a function for Expectation-Maximization algorithm
```{r}
# function for E-M algorithm
emf<-function(xx = vv, # a vector with the data
              m = 2, # number of clustres
              method = "crisp", # method used for parameters estimation in each class; options: "crisp" or "soft"
              itera = 10, # number of iterations
              eps = 0.01, # the accepted error to converge (e)
              params = "random", # initial parameters, if "random" 'm' values from 'xx' are randomly select, or we can input a vector with 'm' values
              main="Glucose level", # histogram title
              seed = 0 # value for set.seed()
              ){
  medias=NA
  desvpadrao=NA
  if (params == "random") {
  set.seed(seed)
  param.mean<-sample(xx,m)
  param.sd<-rep(sd(xx)/m,m)
  } else {
    param.mean<-params
    param.sd<-rep(sd(xx)/m,m)
  }
  
  for (k in 1:itera){
    densities<-sapply(1:m, 
                      function(i){dnorm(xx,
                                        mean=param.mean[i],
                                        sd=param.sd[i])})
    relevance<-t(apply(densities,1,function(l){l/sum(l)}))
    
    attrib<-t(apply(relevance,1,
                    function(l){replace(rep(0,m),which.max(l)[1],1)}))
    
    # Re-estimates the parameters of the probability distribution of each class
    if (method == "crisp") {
      ## M-step (alternative 1: crisp)
      param.mean.new<-sapply(1:m,
                             function(j){mean(xx[attrib[,j]==1])})
      param.sd.new<-sapply(1:m,
                           function(j){sqrt(sum(relevance[,j]*(xx-param.mean[j])^2)/sum(relevance[,j]))})
    } else {
      ## M-step (alternative 2: soft)
      param.mean.new<-sapply(1:m,
                             function(j){sum(xx*relevance[,j])/sum(relevance[,j])})
      param.sd.new<-sapply(1:m,
                           function(j){sqrt(sum(relevance[,j]*(xx-param.mean[j])^2)/sum(relevance[,j]))})
      
    }
    
    # convergence the maximum number of iterations (k).
    cat("iteration", k, "evolution:", 
        dif<-sum(abs(param.mean.new - param.mean)), "\n")
    if(dif < eps) break # the accepted error to converge (e)
    param.mean<-param.mean.new
    param.sd<-param.sd.new
  }
  
  ## ploting the histogram and density curve for the vector of data
  hist(xx, density = 20,
       main=main,
       xlab="", freq=F)
  lines(density(xx),
        lwd=2,col="black")
  
  ## save means, std deviations and classes' attributes on a list
  result<-list(medias = param.mean,
               desvpadrao = param.sd,
               classe = as.data.frame(attrib))

}
```

### 2 clusters density curves
```{r}
# reading the data
breast<-read.csv("2.UploadedData/dataR2.csv")

par(mfrow=c(1,2))

# run the EM function for Glucose with 2 clusters and max 20 iterations
res<-emf(xx = breast$Glucose, itera = 20, m=2, method = "crisp", main = "Glucose values", seed = 0)

# add density curves for each class
lines(density(rnorm(100, res$medias[1], res$desvpadrao[1]), bw = 12),lwd=2,col=5)
lines(density(rnorm(100, res$medias[2], res$desvpadrao[2]), bw = 12),lwd=2,col=6)

## column with 2 classes
attach(res$classe)
classe2<-ifelse(V1 == 1, 5 , 6) # create 'm' classes
detach()

# Plot the values in which each point has a color corresponding to the class to which it belongs
plot(1:length(breast$Glucose), breast$Glucose, col=classe2, pch=20,
     xlab="index", ylab="Glucose values",
     main = " Glucose colored by class (m=2)")
```

### 3 clusters density curves
```{r}
par(mfrow=c(1,2))

# run the EM function for Glucose with 3 clusters and max 20 iterations
res<-emf(xx = breast$Glucose, itera = 20, m=3, method = "crisp", main = "Glucose values", seed = 0)

# add density curves for each class
lines(density(rnorm(100, res$medias[1], res$desvpadrao[1]), bw = 12),lwd=2,col=5)
lines(density(rnorm(100, res$medias[2], res$desvpadrao[2]), bw = 12),lwd=2,col=6)
lines(density(rnorm(100, res$medias[3], res$desvpadrao[3]), bw = 12),lwd=2,col=7)

## column with 3 classes
attach(res$classe)
classe3<-ifelse(V1 == 1, 5,
               ifelse(V2 == 1, 6,
                      7)) # create 'm' classes
detach()

plot(1:length(breast$Glucose), breast$Glucose, col=classe3, pch=20,
     xlab="index", ylab="Glucose values",
     main = " Glucose colored by class (m=3)")
```

