## assignment Second Lead from ECG

Assignment from Lesson 3

2020-02-29

Consider the file A0001.mat from the PhysioNet Challenge https://physionetchallenges.github.io/2020/. Using R:

a) Plot the histogram of all the 12 ECG leads with the respective density curve.

b) Using the second lead inverted, apply the expectation-maximization algorithm with 2 and 3 latent classes.

c) Plot the densities for each of the latent classes.

d) Which point in the ECG belongs to each latent class? Plot the ECG in which each point has a color corresponding to the class to which it belongs.

e) Use the difference between the averages of class distribution as the convergence criterion.

f) Generate 1000 data points at random according to a single distribution fitted to the original data.

g) Generate 1000 data points at random according to a mixture of distributions fitted using one of the previous EM computed in line b.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
```

### Defining a function for expectation-maximization (EM) algorithm
```{r}
# function for Expectation-Maximization algorithm
emf<-function(xx = vv, # a vector with the data
              m = 2, # number of clusters
              method = "crisp", # method used for parameters estimation in each class; options: "crisp" or "soft"
              itera = 10, # number of iterations
              eps = 0.01, # the accepted error to converge (e)
              params = "random", # initial parameters, if "random" 'm' values from 'xx' are randomly select, or we can input a vector with 'm' values
              main="ECG", # histogram title
              seed = 0 # value for set.seed()
){
  medias=NA
  desvpadrao=NA
  if (params == "random") {
    set.seed(seed)
    param.mean<-sample(xx,m)
    param.sd<-rep(sd(xx)/m,m)
  } else {
    param.mean<-params
    param.sd<-rep(sd(xx)/m,m)
  }
  
  for (k in 1:itera){
    densities<-sapply(1:m, 
                      function(i){dnorm(xx,
                                        mean=param.mean[i],
                                        sd=param.sd[i])})
    relevance<-t(apply(densities,1,function(l){l/sum(l)}))
    
    attrib<-t(apply(relevance,1,
                    function(l){replace(rep(0,m),which.max(l)[1],1)}))
    
    # Re-estimates the parameters of the probability distribution of each class
    if (method == "crisp") {
      ## M-step (alternative 1: crisp)
      param.mean.new<-sapply(1:m,
                             function(j){mean(xx[attrib[,j]==1])})
      param.sd.new<-sapply(1:m,
                           function(j){sqrt(sum(relevance[,j]*(xx-param.mean[j])^2)/sum(relevance[,j]))})
    } else {
      ## M-step (alternative 2: soft)
      param.mean.new<-sapply(1:m,
                             function(j){sum(xx*relevance[,j])/sum(relevance[,j])})
      param.sd.new<-sapply(1:m,
                           function(j){sqrt(sum(relevance[,j]*(xx-param.mean[j])^2)/sum(relevance[,j]))})
      
    }
    
    # convergence the maximum number of iterations (k).
    cat("iteration", k, "evolution:", 
        dif<-sum(abs(param.mean.new - param.mean)), "\n") # using the difference between the averages of class distribution as the convergence criterion.
    if(dif < eps) break # the accepted error to converge (e)
    param.mean<-param.mean.new
    param.sd<-param.sd.new
  }
  
  ## ploting the histogram and density curve for the vector of data
  hist(xx, density = 20,
       main=main,
       xlab="", freq=F, ylim = c(0,0.005))
  lines(density(xx),
        lwd=2,col="black")
  
  # plot(xx, col=apply(-densities,1,which.max), 
  #      xlab = "Index", 
  #      ylab = "ECG Values", main = "Colored by class") #plot colours according to vector densities
  
  
  ## save means, std deviations and classes' attributes on a list
  result<-list(medias = param.mean,
               desvpadrao = param.sd,
               classe = as.data.frame(attrib))
  
}
```

### reading the data and ploting the leads distribution
```{r}
library(R.matlab)

# after unzip the files, I took the first file 
fA1<-readMat("2.UploadedData/A0001.mat")$val
fA1<-t(fA1)

par(mfrow=c(3,4))

# Ploting the histogram of all the 12 ECG leads with the respective density curve.
for(i in 1:12){
hist(fA1[,i], density = 10,
     main=paste("ECG lead",i, sep = " "),
     freq=F,
     ylim=c(0,max(density(fA1[,i]*(-1))[["y"]])),
     xlab="ECG values")
lines(density(fA1[,i]),
      lwd=2,col="black")
}
```

### EM algorithm for 2 latent classes
```{r}
par(mfrow=c(1,1))

# expectation-maximization algorithm with 2 latent classes and a maximum 50 iterations.
res<-emf(xx = fA1[,2]*(-1), itera = 50, m=2, method = "crisp", main = "second lead inverted", seed = 10)

# adding density curves from the 2 clusters
lines(density(rnorm(1000, res$medias[1], res$desvpadrao[1]), bw = 40),lwd=2,col=3)
lines(density(rnorm(1000, res$medias[2], res$desvpadrao[2]), bw = 40),lwd=2,col=4)

## column with 2 classes to classify data from each cluster
attach(res$classe)
classe2<-ifelse(V1 == 1, 3 ,4) # create 'm' classes
detach()

# ploting ECG inverted points according with the latent class it belongs to
plot(1:length(fA1[,2]), fA1[,2]*(-1), col=classe2, pch=20,
     xlab="index", ylab="2nd lead ECG",
     main = " ECG colored by class (m=2)")
```

### EM algorithm for 3 latent classes
```{r}
par(mfrow=c(1,1))

# expectation-maximization algorithm with 3 latent classes and a maximum 50 iterations.
res<-emf(xx = fA1[,2]*(-1), itera = 50, m=3, method = "crisp", main = "second lead inverted", seed = 10)
# adding density curves from the 3 clusters
lines(density(rnorm(1000, res$medias[1], res$desvpadrao[1]), bw = 40),lwd=2,col=3)
lines(density(rnorm(1000, res$medias[2], res$desvpadrao[2]), bw = 40),lwd=2,col=4)
lines(density(rnorm(1000, res$medias[3], res$desvpadrao[3]), bw = 40),lwd=2,col=5)

## column with 3 classes to classify data from each cluster
attach(res$classe)
classe3<-ifelse(V1 == 1, 3 ,
               ifelse(V2 == 1, 4, 5)) # create 'm' classes
detach()

# ploting ECG inverted points according with the latent class it belongs to
plot(1:length(fA1[,2]), fA1[,2]*(-1), col=classe3, pch=20,
     xlab="index", ylab="2nd lead ECG",
     main = " ECG colored by class (m=3)")

```

### Generating 1000 data points at random with bootstrapped, with guassian distribution and EM with 3 clusters.
```{r}
## length of data set generation
n<-1000

# original sample data set
sample<-fA1[,2]*(-1)

# bootstrapped sample data set
sample.obs<-sample(fA1[,2]*(-1), size=n, replace = TRUE)

# gaussian generated data set
sample.simple<-rnorm(n,mean=mean(fA1[,2]*(-1)), sd=sd(fA1[,2]))


# EM model generated data set
#classes<-apply(attrib, 1, which.max) # using attrib define previously, to calculate the number of observations in each class
class.prop<-prop.table(table(classe3)) # calculate percentage of each class
m<-3
set.seed(523)
sample.classes<-sample(1:m, size=n, replace=TRUE, prob=class.prop)
sample.prop<-prop.table(table(sample.classes))
sample.values<-round(unlist(sapply(1:m,
                                   function(j){
                                     rnorm(n*sample.prop[j], # generating number with the proportions of the two classes we identified previously
                                           mean=res$medias[j], # parameters defined previously
                                           sd=res$desvpadrao[j]) # parameters defined previously
                                   })),0)

# Plotting results
par(mfrow=c(2,2))

hist(sample, freq = F, main = "Observed data", ylim = c(0,0.004))
lines(density(sample))

hist(sample.obs, freq = F, main = "data points sampled from observed data", ylim = c(0,0.004))
lines(density(sample.obs))

hist(sample.simple, freq = F, main = "data points generated from sample distribution", ylim = c(0,0.004))
lines(density(sample.simple))

hist(sample.values, freq = F, main = "data points generated from EM (m=3)", ylim = c(0,0.004))
lines(density(sample.values))

```


