## Lesson 3

2020-02-18

### Using ‘rpart’ and the Breast Cancer Coimbra data set

```{r}
# Define the filename
filename <- "2.UploadedData/dataR2.csv"

# Load the CSV file from the local directory
dataset <- read.csv(filename, header=T)
dataset$Classification<-factor(dataset$Classification, levels=1:2, labels=c("Controls","Patients"))
```

```{r}
library(rpart)

# learn tree with Gini impurity
tree.gini<-rpart(Classification~.,data = dataset)

# learn tree with information gain
tree.information<-rpart(Classification~., data=dataset,
                        parms=list(split="information"))
```

Summary Information
```{r}
summary(tree.information)

```

Summary Gini
```{r}
summary(tree.gini)
```

Plot
```{r}
par(mfrow=c(1,2))
plot(tree.information,main="Information")
text(tree.information)
plot(tree.gini, main="Gini")
text(tree.gini)
```

### Random Forest
```{r}
library(caret)
library(pROC)
```

Run algorithms using 3 times 10-fold cross validation
```{r}
# set definitions
metric <- "ROC"

control <- trainControl(method="repeatedcv", number=10,
                        summaryFunction=twoClassSummary, 
                        classProbs=T,
                        savePredictions = TRUE, repeats=3)
```

train Tree and RF
```{r}
set.seed(7)
fit.cart.rcv <- train(Classification ~ ., data=dataset, method="rpart", metric=metric, trControl=control)

set.seed(7)
fit.rf.rcv <- train(Classification ~ ., data=dataset, method="rf", metric=metric, trControl=control)

```

summarize accuracy of models
```{r}
fit.models <- list(rpart=fit.cart.rcv, rf=fit.rf.rcv)
results <- resamples(fit.models)
summary(results)
```

ROC curves for models
```{r}
par(mfrow=c(1,2))
rocs <- lapply(fit.models, function(fit){plot.roc(fit$pred$obs,fit$pred$P, 
                                                  main=paste("3x10-fold-CV",fit$method), 
                                                  debug=F, print.auc=T)})
```

Compare accuracy of models
```{r}
dotplot(results)
```

Inspect models
```{r}
print(fit.cart.rcv)
#getModelInfo(fit.cart.rcv)
#getModelInfo(fit.cart.rcv)$rpart
getModelInfo(fit.cart.rcv)$rpart$parameters
```

Inspect models
```{r}
print(fit.rf.rcv)
#getModelInfo(fit.rf.rcv)
#getModelInfo(fit.rf.rcv)$rf
getModelInfo(fit.rf.rcv)$rf$parameters
```

ROC complexity for models
```{r}
plot(fit.cart.rcv)
plot(fit.rf.rcv)
```

### Improve Random Forest
```{r}
myGrid<-expand.grid(mtry=1:9) # numero de variaveis/colunas a considerar na RF

set.seed(7)
fit.rf.rcv.tune <- train(Classification ~ ., data=dataset, 
                         method="rf", metric=metric, 
                         trControl=control,
                         tuneGrid = myGrid)
```

summarize accuracy of models
```{r}
fit.models<-list(rpart=fit.cart.rcv,rf = fit.rf.rcv, rf.tune=fit.rf.rcv.tune)

results<-resamples(fit.models)
summary(results)
```

Roc curves for models
```{r}
par(mfrow=c(1,3))
rocs<-lapply(fit.models,
             function(fit){plot.roc(fit$pred$obs,
                                    fit$pred$P,
                                    main=paste("3x10-fold CV-",fit$method),debug=F,print.auc=T)})
```

Compare accuracy of models
```{r}
dotplot(results)
```

Inspect models
```{r}
getModelInfo(fit.rf.rcv.tune)$rf$parameters
```

