## Assignment Trees 

2020-02-21

Compare a decision tree with a random forest in the Cervical Cancer (Risk Factors) data set (available from UCI repository), trying to accurately classify Dx.Cancer

```{r }
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

### load data

```{r}
library(tidyverse)
library(caret)
library(rpart)
library(rpart.plot)
library(pROC)

data <- read.csv("2.UploadedData/risk_factors_cervical_cancer.csv", na.strings = "?")

## select columns 0/1
colMax <- function(data) sapply(data, max, na.rm = TRUE)

cols<-colMax(data) %>% data.frame(max=.) %>% rownames_to_column("cols") %>% 
  filter(max==1) %>% .$cols

# factorize columns 0/1
data<-data %>% mutate_at(cols, ~factor(.,levels = 0:1, labels = c("no","yes")))

# excluded columns only with zeros and Dxs
data<-data %>% select(-STDs.AIDS,-STDs.cervical.condylomatosis,
                      -Dx.CIN, -Dx.HPV, -Dx)
```

summarize the data
```{r}
summary(data)
table(data$Dx.Cancer)
```

missing data
```{r}
library(Amelia)
missmap(data)
```

new data
```{r}
# variables with a low number of missings
vars<-c("Age","Hinselmann","Schiller","Citology","Biopsy","Number.of.sexual.partners"
        ,"Smokes..packs.year.","Smokes..years.","Smokes", "First.sexual.intercourse",
        "STDs..Number.of.diagnosis", "Dx.Cancer")
data.new<-data %>% select(vars)
```

### Classification tree
```{r}
model1 <- rpart(Dx.Cancer ~ ., data = data.new,
               method="class", control =rpart.control(minsplit =1,minbucket=3, cp=0)
               )
par(xpd = NA) # otherwise on some devices the text is clipped
plot(model1)
text(model1, digits = 3)

print(model1, digits = 2)

```

```{r}
## accuracy on train data
predicted.classes <- model1 %>% 
  predict(data.new, type = "class")

# Compute model accuracy
mean(predicted.classes == data.new$Dx.Cancer)
```

train model Tree with repeated 3*10-fold CV
```{r}
# set definitions
metric <- "ROC"

control <- trainControl(method="repeatedcv", number=10,
                        summaryFunction=twoClassSummary, 
                        classProbs=T,
                        savePredictions = TRUE, repeats=3)

# regression tree with repeated 3*10-fold CV
set.seed(1)
model2 <- train(
  Dx.Cancer ~ ., 
  data = data.new, 
  method = "rpart",
  trControl = control,
  metric = metric,
  na.action=na.exclude
  #tuneLength = 10
)

```

Accuracy of the model2
```{r}
predicted.prob <- model2 %>% 
  predict(data, type = "prob")
pred.class<-ifelse(predicted.prob$yes < 0.5, "no", "yes")
mean(pred.class == data.new[complete.cases(data.new),]$Dx.Cancer)
```

### Random forest with repeated 3*10-fold CV
```{r}
set.seed(1)
model3 <- train(
  Dx.Cancer ~ ., 
  data = data.new, 
  method = "rf",
  trControl = control,
  metric = metric,
  na.action=na.exclude
)
```

compare models
```{r}
fit.models <- list(rpart=model2, rf=model3)
results <- resamples(fit.models)
summary(results)
```

plot ROC
```{r}
par(mfrow=c(1,2))
rocs <- lapply(fit.models, function(fit){plot.roc(fit$pred$obs,fit$pred$yes, 
                                                  main=paste("3x10-fold-CV",fit$method), 
                                                  debug=F, print.auc=T)})
```

compare accuracy  of models
```{r}
dotplot(results)
```

Inspect model Tree
```{r}
print(model2)
```

inspect model Random Forest
```{r}
print(model3)
```

ROC / complexity for models
Tree
```{r}
#plot(model2)
```

RF
```{r}
plot(model3)
```

### Improve Random Forest
```{r}
myGrid<-expand.grid(mtry=1:11) # number of independent variables

set.seed(1)
model3.tune <- train(Dx.Cancer ~ ., data=data.new, 
                         method="rf", metric=metric, 
                         trControl=control,
                        na.action=na.exclude,
                         tuneGrid = myGrid)
```

compare models
```{r}
fit.models<-list(rpart=model2,rf=model3,rf.tune=model3.tune)
results<-resamples(fit.models)
summary(results)
```

ROC curves
```{r}
par(mfrow=c(1,3))
rocs<-lapply(fit.models,
             function(fit){plot.roc(fit$pred$obs,
                                    fit$pred$yes,
                                    main=paste("3x10-fold CV-",fit$method),debug=F,print.auc=T)})
```

ROC, sens and spec
```{r}
dotplot(results)
```


http://www.sthda.com/english/articles/35-statistical-machine-learning-essentials/141-cart-model-decision-tree-essentials
https://bradleyboehmke.github.io/HOML/DT.html
https://www.machinelearningplus.com/machine-learning/logistic-regression-tutorial-examples-r/

