---
title: "COMSTATS - lesson5"
author: "Bruno A Lima"
date: "2020-03-16"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---


```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

## Estimate the value of $\pi$ through a Monte Carlo simulation

star info:

- the area of a square of side $2R$: $A_s = 4R^2$
- the area of a circle with radius $R: A_c = \pi R^2$
- if we know the areas of both forms, $A_s$ and $A_c$, then

$A_s/4 = A_c/ \pi$ or $\pi = 4A_c/A_s$

- a point $p$ is inside a circle of radius $R$ if $D(0,p) < R$ 

### function to compute euclidean distances
```{r}
euclid <- function(a, b) sqrt (sum((a-b)^2))
```

### function to compute pi from simulation
```{r}
pi.run <- function(nsamples, square) {
  # compute the radius as half of a side
  r <- (square[2] - square[1]) / 2
  # center of the circle
  zero <- rep((square[2] + square[1]) / 2, 2)
  # data frame with points x and y, uniforme distributed between begin and end of the square
  xy <- data.frame(xx=runif(nsamples, min=square[1], max=square[2]), 
                   yy=runif(nsamples, min=square[1], max=square[2]))
  # add column with T/F if distance of xy to zero lower than r (i.e xy is inside the circle)
  xy <- cbind(xy, inside=apply(xy, 1, function(p) {euclid(p, zero)}) <= r)
  # add column with the cumulative Trues divided by number of the iteration 
  xy <- cbind(xy, prop = cumsum(xy$inside) / 1:nsamples)
  # compute the aproximation of pi at each iteration
  xy <- cbind(xy, pi = 4 * xy$prop)
  xy
}
```

### prepare plot with a empty square 
```{r}
preparePlot <- function(square) {
  plot(NULL, xlim=square, ylim=square, axes=F, xlab="x", ylab="y")
  axis(1)
  axis(2)
  axis(3)
  axis(4)
  lines(x=c(square[1],square[1]), y=c(square[1],square[2]))
  lines(x=c(square[1],square[2]), y=c(square[1],square[1]))
  lines(x=c(square[1],square[2]), y=c(square[2],square[2]))
  lines(x=c(square[2],square[2]), y=c(square[1],square[2]))
}
```

### define the sizes of a square centered at zero (i.e a vector with simetric values)
```{r}
square <- c(-1,1)
```

### define the number of iterations
```{r}
nsamples <- 100000 #1e5
```

### par(mfrow=c(1,3))
```{r}
set.seed(523)
xy <- pi.run(nsamples,square)
```

### plot the circle
```{r}
preparePlot(square)
points(x=xy$xx, y=xy$yy, col=ifelse(xy$inside,2,1), pch=".", cex = 2)
```

### plot the convergence to pi
```{r}
plot(xy$pi, axes=F, main="PI Estimate Evolution - Single Monte Carlo", xlab="Iteration", ylab="pi", type="l", ylim=mean(xy$pi)+3*c(-sd(xy$pi),sd(xy$pi)))
axis(1)
axis(2)
lines(c(0,nsamples), rep(pi,2), lty="dashed")
```

### define number of Monte Carlo runs
```{r}
mc.runs <- 1e3 / 2
```

### compute pi the defined number of runs, keeping the last value of the column pi from pi.run each time
```{r}
pi.est <- unlist(lapply((1:mc.runs), function(a) pi.run(nsamples,square)$pi[nsamples]))

```

### plot an histogram with obtained aproximations of pi
```{r}
hist(pi.est, main=paste("Distribution of PI Estimates (",mc.runs, " runs of ", nsamples," samples)",sep=""))
text(x=3.155, y=250, labels=paste("mu = ", round(mean(pi.est),5),sep=""))
```

### and now FRANz's solution:

define a funcion to compute pi
```{r}
random_area<-function(n=1e5, radius = 0.5){
  x<-runif(n, -radius, radius)
  y<-runif(n, -radius, radius)
  
  dist<-sqrt(x^2+y^2) < radius
  
  res<-sum(dist)/n * 4
  
  return(res)
}
```

do the MC and plot the histogram with the results
```{r}
trails<-NULL

for (i in 1:1000){
  trails<-c(trails, random_area())
}

hist(trails)
```

### and also Seref's solution:

define a function
```{r}
simulation = function(long){
  c = rep(0,long)
  numberIn = 0
  for(i in 1:long){
    x = runif(2,-1,1)
    if(sqrt(x[1]*x[1]+x[2]*x[2])<=1){
      numberIn = numberIn + 1
    }
    prop = numberIn / i
    piHat = prop * 4
    c[i] = piHat
  }
  return(c)
}
```

and the Monte Carlo part in a plot
```{r}
size = 10000
res = simulation(size)
ini = 1
plot(res[ini:size], type = "l")
lines(rep(pi,size)[ini:size],col = "red")
```

