---
title: "LEARN assignment Trees and RF (up sampling)"
author: "Bruno A Lima"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---

## Assignment Trees - Up Sampling

Compare a decision tree with a random forest in the Cervical Cancer (Risk Factors) data set (available from UCI repository), trying to accurately classify Dx.Cancer

In this excercise we have a disparity in the frequencies of the observed classes we want to predict.  For resolving such a class imbalance we can up sampling our data as described elsewhere:
https://topepo.github.io/caret/subsampling-for-class-imbalances.html#subsampling-techniques

```{r }
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

### load data

```{r}
library(tidyverse)
library(caret)
library(rpart)
library(rpart.plot)
library(pROC)
data <- read.csv("risk_factors_cervical_cancer.csv", na.strings = "?")
## select columns 0/1
colMax <- function(data) sapply(data, max, na.rm = TRUE)
cols<-colMax(data) %>% data.frame(max=.) %>% rownames_to_column("cols") %>% 
  filter(max==1) %>% .$cols
# factorize columns 0/1
data<-data %>% mutate_at(cols, ~factor(.,levels = 0:1, labels = c("no","yes")))
# excluded columns only with zeros and Dxs
data<-data %>% select(-STDs.AIDS,-STDs.cervical.condylomatosis,
                      -Dx.CIN, -Dx.HPV, -Dx)
```

summarize the data
```{r}
summary(data)
table(data$Dx.Cancer)
```

### Up sample data
```{r}
# upsample data
set.seed(1)
up_data <- upSample(x = data[, !colnames(data) %in% "Dx.Cancer"],
                     y = data$Dx.Cancer)
# sumarize it
# now the dependent variable is named 'Class'
table(up_data$Class)
```


missing data
```{r}
library(Amelia)
# missings on original data
missmap(data)
```

new data
```{r}
# select only variable with low number of missings
vars<-c("Age","Hinselmann","Schiller","Citology","Biopsy","Number.of.sexual.partners"
        ,"Smokes..packs.year.","Smokes..years.","Smokes", "First.sexual.intercourse",
        "STDs..Number.of.diagnosis", "Class")
up_data<-up_data %>% select(vars)
```


### Classification tree by two methods
```{r}
# learn tree with Gini impurity
tree.gini<-rpart(Class~.,data = up_data, cp = 0.1)
# learn tree with information gain
tree.information<-rpart(Class~., data=up_data,
                        parms=list(split="information"))
par(mfrow=c(1,2))
plot(tree.information,main="Information")
text(tree.information)
plot(tree.gini, main="Gini")
text(tree.gini)
```

### train models CART and RF with repeated 3*10-fold CV
```{r}
# 
# set definitions
metric <- "ROC"
control <- trainControl(method="repeatedcv", number=10,
                        summaryFunction=twoClassSummary, 
                        classProbs=T,
                        savePredictions = TRUE, repeats=3)
set.seed(7)
fit.cart.rcv <- train(Class ~ ., data=up_data, method="rpart", metric=metric, trControl=control,
                      na.action=na.exclude)
set.seed(7)
fit.rf.rcv <- train(Class ~ ., data=up_data, method="rf", metric=metric, trControl=control,
                    na.action=na.exclude)
```

compare CART with RF 
```{r}
fit.models <- list(rpart=fit.cart.rcv, rf=fit.rf.rcv)
results <- resamples(fit.models)
summary(results)
par(mfrow=c(1,2))
rocs <- lapply(fit.models, function(fit){plot.roc(fit$pred$obs,fit$pred$yes, 
                                                  main=paste("3x10-fold-CV",fit$method), 
                                                  debug=F, print.auc=T)})
dotplot(results)
print(fit.cart.rcv)
print(fit.rf.rcv)
plot(fit.cart.rcv)
plot(fit.rf.rcv)
```

tunning RF model
```{r}
myGrid<-expand.grid(mtry=1:11) # number of variables  when constructing the RF
set.seed(7)
fit.rf.rcv.tune <- train(Class ~ ., data=up_data, 
                         method="rf", metric=metric, 
                         trControl=control,
                         tuneGrid = myGrid,
                         na.action=na.exclude)
```

compare the 3 models
```{r}
fit.models<-list(rpart=fit.cart.rcv,rf = fit.rf.rcv, rf.tune=fit.rf.rcv.tune)
results<-resamples(fit.models)
summary(results)
```

ROCs from models
```{r}
par(mfrow=c(1,3))
rocs<-lapply(fit.models,
             function(fit){plot.roc(fit$pred$obs,
                                    fit$pred$yes,
                                    main=paste("3x10-fold CV-",fit$method),debug=F,print.auc=T)})
```

Compare accuracy of models
```{r}
dotplot(results)
```

